---
bibliography: Paperpile.bib
csl: nature.csl
geometry: margin=2.5cm
linestretch: 2
mainfont: mathptmx
output:
  html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    theme: sandstone
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: no
    pandoc_args:
    - +RTS
    - -K64m
    - -RTS
  word_document:
    fig_caption: yes
    fig_width: 7.5
---

<!-- ## Install and source necessary packages-->
```{r, echo=FALSE, include=FALSE}


  # install.packages('knitcitations')
  # install.packages('bibtex')
  # install.packages('qdap')

  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries=FALSE)
  library(bibtex)
  
  # source('http://bioconductor.org/biocLite.R')
  # biocLite('phyloseq')
  # packageVersion('phyloseq')
  library(phyloseq)
  # library("qdap") # to be able to use word count
  library("plyr")
  require("dplyr")
  require("mosaic")
```
\pagebreak

TITLE {-}

Authors: Danielle C. Claar $^1$, Kristina L. Tietjen $^1$, Ruth D. Gates $^2$, Julia K. Baum $^1$

Institute: $^1$ Department of Biology, University of Victoria, PO BOX 1700 Station CSC, Victoria, British Columbia, V8W 2Y2, Canada
$^2$ Hawaii Institute of Marine Biology, 46-007 Lilipuna Road, Kaneohe, HI 96744, USA

Corresponding Author: Danielle C. Claar, Tel: (208) 250-0161, Email: dclaar@uvic.ca

Keywords: coral bleaching, heat stress, climate change, Symbiodinium

\pagebreak 

# Summary {-} 
*Let's leave this for now and come back to it once we have rest of manuscript honed* Coral reefs, which already live on the edge of their thermal tolerance [@Sampayo2016-vd], are under acute threat from ocean warming [@Hughes2003-aj; @Hoegh-Guldberg2007-fh; @Baker2008-ky]. Corals live in symbiosis with an extraordinarily diverse genus of photosynthetic dinoflagellates (Symbiodinium spp.; [@Muscatine1977-pn; @Rowan1992-lg]). The symbiotic association and diversity of various taxa of *Symbiodinium* can be flexible over time [@Baker2003-ks; @Little2004-tm], but see [@Thornhill2006-cs; @Thornhill2009-jh], and individual *Symbidoinium* taxa can range from parasites to mutualists in their interaction with their coral host [@Lesser2013-dj]. Warming causes the breakdown of coral symbiosis, causing coral "bleaching" when symbionts are expelled and the white coral skeleton is visible through the coral tissue [@Brown1997-mf]. Coral bleaching can lead to mortality, although corals can regain their symbionts after heat stress has abated [@Douglas2003-nr; @Stat2009-qq]. The 2015/16 El Niño is the worst pulse warming event on record in terms of severity and longevity [@Eakin:2016vf; @Heron2016-am], yet despite massive coral mortality, some corals show resilience to this extreme event (@Bauminprep). Here, we track coral symbioses and survival at the epicenter of this bleaching event (Kiritimati, Central Pacific), and show, contrary to our current paradigm of coral bleaching and recovery dynamics, that some corals have the capacity to re-establish symbiosis before heat stress subsides. Furthermore, we demonstrate potential mechanisms for coral survival and recovery, including the lack of preferential symbiont expulsion, and the effect of local human disturbance on  pre-bleaching symbiont community structure and the probability of coral survival. Together, these results show the potential for reef corals to survive extreme warming events, providing tentative hope for the survival of corals in the Anthropocene.

\pagebreak

# Main Text {-}
```{r, echo=FALSE, include=FALSE}
citep("Article")
```
NEED INTRO SENTENCE TO HOOK READERS. The symbiosis between coral and their single-celled dinoflagellate symbionts, *Symbiodinium*, is the foundation of reef ecosystems, and a critical element of reef resilience [@Van_Oppen2006-qf; Muller-Parker2015-sd]. The coral holobiont responds to environmental conditions, and is the unit that interacts with the broader reef community [@Gates2011-zy], supporting reef diversity and function across taxa *(or at a global scale?)*. There is much genetic, functional, and response diversity within the *Symbiodinium* genus. Although *Symbiodinium* is currently classified as a single genus, it contains diversity similar to diversity found within other dinoflagellate orders [@Rowan1992-lg], and is divided into nine clades and hundreds of types [@Van2005-di; @Franklin2012-sz; @Tonk2013-pf]. There is much debate about species classifications within *Symbiodinium*, which has hampered species naming within *Symbiodinium* [@Stat2012-gw; but see @LaJeunesse2001-hf) although *Symbiodinium* types are generally considered putative species [@Pochon2010-jm. *Symbiodinium* types have distinct geographic distributions, host associations, and environmental optima [@Fabina2012-mm]. Furthermore, total and relative abundance of *Symbiodinium* can vary among coral colonies, across environmental gradients, and over time [@Fitt2000-ex], with increased *Symbiodinium* abundance leading to increased environmental sensitivity and bleaching risk [@Cunning2012-um]. There are functional differences between *Symbiodinium* clades [@Stat2008-hk], and *Symbiodinium* associations can range from mutualistic to neutral to parasitic based on *Symbiodinium* type as well as environmental conditions [@Lesser2013-dj]. Recent advances in next-generation sequencing techniques have revealed cryptic genetic diversity within symbiotic *Symbiodinium* [@Quigley2014-zj; @Arif2014-kx; @Green2014-az], and has allowed for long-term genetic and ecological comparisons of symbiont community structure [@Edmunds_undated-fd].  
  
Corals exhibit varying levels of symbiotic flexibility, but this flexibility comes with functional tradeoffs. Corals that host flexible symbioses (generalists) may be more sensitive to environmental perturbations than those with intimate symbioses (specialists) [@Putnam2012-om]. The adaptive bleaching hypothesis suggests that corals bleach in order to expel environmentally sub-optimal symbionts, followed by switching (picking up new symbionts from the environment) or shuffling (an internal change in dominant symbiont type or overall symbiont community structure) [@Buddemeier2004-se; @Buddemeier1993-sx; @Baker2001-vc; @Baker2003-ks]. There is ample evidence for *Symbiodinium* shuffling (Rowan 2004), and a recent study showed evidence for *Symbiodinium* switching [@Boulotte2016-dy]. However, what remains unclear is if and how frequently bleaching events can actually be considered adaptive. Changes in photosynthetic efficiency during bleaching as well as bleaching resistance have been shown to correspond to distinct *Symbiodinium* phylotypes [@Kemp2014-xj]. Clade D *Symbiodinium* are proported to have an enhanced thermal tolerance [@Stat2010-zg], and repopulation of a coral host with clade D symbionts after a bleaching event is proposed to be a survival mechanism [@Berkelmans2006-rf; @Mieog2007-yy; @Silverstein2012-tm]. A history of thermal stress increased the prevalence of clade D *Symbiodinium* in a generalist coral species, but did not instigate similar changes in two specialist coral species [@Stat2013-qp]. Although the prevalence of clade D *Symbiodinium* increases during thermal stress and may increase thermal tolerance [@Berkelmans2006-rf], corals that house clade D symbionts may have slower growth rates [@Little2004-tm] or lower energy storage [@Jones2011-nf]. Furthermore, functional differences exist not only at the clade level, but are present among types within a single clade [@Sampayo2008-tw].
  
The current paradigm of coral bleaching and recovery states that the stress must cease for coral to regain their symbiosis. Coral bleaching is the loss of obligate symbionts (Symbiodinium) from the coral tissue [@Gates1992-ew; @Douglas2003-nr]. Thermal stress is the primary cause for coral bleaching, and can cause not only the breakdown of coral symbioses, but also cause coral mortality [@Hoegh-Guldberg1999-rb; CITE]. Thermal stress can be exacerbated by other environmental stressors (Cooper et al 2011, Béraud et al 2013, Maina et al 2008), and in turn, exacerbates ocean acidification (Gibbin et al 2015). The current paradigm of coral bleaching and resilience is that as environmental stress (such as warming) increases, corals begin to bleach. Extreme or long-lasting warming causes a complete breakdown of the coral symbioses, leading to expulsion of all (or nearly all) *Symbiodinium* from the coral host tissue. It has been shown that during bleaching, there is a window for recovery, that is, a certain amount of time during which the warming must cease and conditions must return to normal so that the coral can regain its symbionts. If the window for recovery passes without amelioration of the environmental conditions, the coral will starve and die. (Cunning et al 2016, Putnam et al 2017). Here we show that despite unprecedented heat stress, some corals exhibited resilience and survived. Survival through such an extreme heat event provides an exceptional opportunity to understand how some corals can withstand intense heat stress, and how corals in general might survive long-term warming. Remarkably, we find that some coral colonies were able to survive this prolonged heat stress by regaining their symbionts while temperatures were still elevated.
  
  
Global coral bleaching is increasing, and the 2014-2017 event caused a catastrophic loss of corals around the globe. There was up to 95% mortality in some regions during the 1997/1998 El Niño event (Glynn 1993). The 2014-2017 global coral bleaching event caused coral bleaching across the world's oceans (Eakin 2016, Normile 2016), with up to 75% bleaching on some reefs in Hawaii, and at least some level of bleaching across 93% of the Great Barrier Reef (Minton et al 2015, GBRMPA 2016). The 2015-2016 El Niño, superimposed on nearly-ubiquitous tropical ocean warming, instigated the third global coral bleaching event (@Eakin:2016vf). Our study location, Kiritimati Atoll (Christmas Island, Kiribati, Central Equatorial Pacific, Coordinates: 2, -157.4), was at the epicenter of this extreme El Niño event. Thermal anomalies were severe on Kiritimati, rapidly exceeding NOAA Coral Reef Watch's Coral Bleaching Alert Level 1 (4 Degree Heating Weeks, DHW, a metric of cumulative thermal stress) and Alert Level 2 (8 DHW) thresholds, reaching an unprecedented (@Hoegh-Guldberg2011-sl) 25.7 DHW over a year-long bleaching event, demolishing most of the reef (@Baum_inprep). Despite these staggering losses, some corals have the capacity to be resilient to these increasingly frequent mass-bleaching events (Hughes et al 2017). Here, we assess coral symbiosis and survival during the massive 2015/2016 El Niño event. We tagged, sampled, and photographed the same coral colonies before, during, and immediately after the El Niño event. We assessed bleaching condition and survival for each coral colony, and used Illumina MiSeq ITS2 amplicon sequencing and 97% *de novo* OTU clustering to evaluate changes in *Symbiodinium* community structure. To investigate mechanisms underlying the ability of these corals to not only survive a year of continuous heat stress, but to recover in the interim, we assessed the relationship between human disturbance, pre-bleaching *Symbiodinium* community structure, and coral survival, as well as the timing of *Symbiodinium* community shifts throughout this El Niño event. We document, for the first time, corals that were able to visually recover from bleaching, and to regain their *Symbiodinium* communities during the course of an extreme heat stress event. These corals (family Faviidae; *Platygyra* sp. and *Favites* sp.) were bleached within two months of the onset of warming, but had visibly recovered after 10 consecutive months of intense warming (Fig. 1).

```{r, echo=FALSE, warning=FALSE, eval=TRUE,include=FALSE}
source("../figures/Figure1.R")
```
![Figure 1](../figures/Figure1.tiff)


The "transient microbiome" (of rare *Symbiodinium* types) assembled by environmental anomalies can undergo rapid changes (Putnam et al 2017), providing symbiotic stochasticity which may build or weaken a coral's capacity for resilience. While corals have been shown to change symbiotic partners during a bleaching event (Chen et al 2005, Jones et al 2008), there is often a quick return to pre-bleaching *Symbiodinium* communities after recovery (Thornhill 2005, LaJeunesse et al 2010), and persistence of stress-related changes to *Symbiodinium* community structure may require sustained environmental pressure (Baird et al 2007). Corals commonly host background *Symbiodinium* types in low levels (Correa et al 2009), but sub-dominant *Symbiodinium* communities are often unstable (Coffroth et al 2010). The importance of rare *Symbiodinium* types is currently under debate, and these rare types may be commensal (pass through coral's holobiont with no harm or gain for either partner), parasitic ("cheaters", or symbionts that take more than they give), or mutualistic (symbionts which support host function) (Parkinson et al 2015). Some research suggests that low-abundance *Symbiodinium* types have minimal functional significance to corals (Lee et al 2016), while other evidence supports the idea that the rare *Symbiodinium* biosphere is important for corals' response to climate change (Boulotte et al 2016). However, shifts in *Symbiodinium* community diversity may still have a larger influence on coral resilience than the evolution of symbiont thermal tolerance (Baskett et al 2010). In other systems, other rare microbial species have been demonstrated to be disproportionally important to maintaining functional processes during environmental change (Shade et al 2014). We show that after two months of heat stress, fully-bleached corals retained approximately the same *Symbiodinium* community as they had before the bleaching event. This suggests that a wholesale breakdown of symbiosis occurred in bleached corals during this event, indicating a lack of preferential symbiont expulsion or exodus. Furthermore, we demonstrate that symbionts present in even very low abundances can play a critical role in coral survival and recovery. Some coral colonies recovered symbiosis with *Symbiodinium* types that were present in only a negligible amount before the bleaching event.
  
Local protection is critical for coral symbiosis and survival. We show that corals living at different levels of local human disturbance had distinct symbiont communities that corresponded tightly to survivorship. This is in contrast to a recent study which concluded that particulate and dissolved nutrients do not reduce coral health at a colony scale (Rocker et al 2017). However, there is increasing evidence for local adaptation in corals (Howells et al 2012, Logan et al 2013, Dixon et al 2015). Our results suggest that some Kiritimati coral species may have the capacity to experience evolutionary rescue, defined as adaptation at a rate that allows an endangered population to survive the rate of environmental change (Orr & Unkless  2014, Carlson 2014). Our results suggest that the capacity for evolutionary rescue is tangibly related to local reef protection. Although massive bleaching events like this one will likely continue to cause catastrophic damage to coral reefs worldwide, mitigating local human disturbance can potentially help protect some coral species against a modest amount of ocean warming. 

# Methods {-}
*The Methods section should be written as concisely as possible but should contain all elements necessary to allow interpretation and replication of the results. As a guideline, Methods sections typically do not exceed 3,000 words. Detailed descriptions of methods already published should be avoided; a reference number can be provided to save space, with any new addition or variation stated. The Methods section should be subdivided by short bold headings referring to methods used and we encourage the inclusion of specific subsections for statistics, reagents and animal models. If further references are included in this section, the numbering should continue from the end of the last reference number in the rest of the paper and the list should accompany the additional Methods at the end of the paper. The Methods section cannot contain figures or tables (essential display items should be included in the Extended Data).*

## Field Sampling 

Kiritimati Atoll (Christmas Island), Kiribati is located in the Central Equatorial Pacific (1.9N 157W), at the center of the El Niño 3.4 region (a region which is used to quantify El Niño presence and strength). During the 2015/2016 El Niño event, Kiritimati experienced 10 months of sustained temperature stress, causing a mass bleaching and mortality event (@BauminPrep). 

## Temperature quantification
Temperature loggers (Sea-Bird 56) were deployed around the island at 10-12m depth from 2011-2016 to measure *in situ* thermal stress.

##Coral Tagging and sampling  
In August/September 2014 colonies of *Platygyra* sp. and *Favites pentagona* were tagged along a 60m transect at 10-12m depth at 15 different sites around the atoll.  A photo was taken of each coral to record colony measurments and characteristics (i.e. bleaching). The tagged coral colonies were resampled twice more before (January/February 2015, April/May 2015), once during (July 2015), and once near the end (March 2016) of the El Niño warming. Some tagged coral colonies were lost due to storm damage, and new coral colonies were tagged to replenish the total number of surveyed colonies. Not all sites were visited during all field seasons, and some site surveys were only partially completed during some field seasons due to inclement weather conditions.

Corals were sampled by... processed like this... Coral tissue samples were preserved in Guanidinium buffer (50% w/v guanidinium isothiocyanate; 50 mM Tris pH 7.6; 10 µM EDTA; 4.2% w/v sarkosyl; 2.1% v/v β-mercaptoethanol) and stored at 4 *degrees* until extraction.

## Pre-processing and sequencing
DNA extraction was performed using a guanidinium-based extraction protocol [@Stat2009-qq; Cunning2017-sc; Cunning2015-mt] with the modification that the DNA pellet was washed with 70% ethanol three times rather than once.

Library Prep- Amy's method of library prep, include cleanup, Illumina Sequencing information (barcodes, etc) for sequencing of ITS2 amplicons (‘itsD’ and ‘its2rev2’ primers from [@Stat2009-qq] 
Illumina MiSeq platform with 2×300 paired-end read chemistry.
HIMB - Amy

ITS2 region - it's annoying, but it's the best we've got right now smith et al 2017

A total of 289 samples were prepared for sequencing, and *XXXX* of these samples were successfully amplified, sequenced, and used in downstream analyses.



## Bioinformatics
We conducted quality filtering of raw reads (in .fastq format) first using iu-filter-quality-bokulich implemented in Illumina-Utils [@Bokulich2013-cm; Eren2013-yg], followed by paired-end sequence merging via iu-merge-pairs (also in Illumina-Utils, [Eren2013-yg]), with a maximum mismatch of three bases between the forward and reverse reads. After quality filtering, sequence processing and identification was performed following all specifications of [@Cunning2017-sc]; chimeric sequences were removed, primers were trimmed, sequences from each sample were clustered independently at 97% similarity using UCLUST [@Edgar2010-zl] implemented in QIIME [@Caporaso2010-yl] and resulting OTUs were collapsed at 100% identity across samples, sequences were aligned using the Needleman-Wunsch global alignment algorithm (Biostrings package, [@Pages2017-ie]) in R [@R_Development_Core_Team2008-sp], and sequences were named using a reference database. 

The Phyloseq package [@McMurdie2013-hf] in R was used to store and analyze OTU tables, taxonomic information, and sample metadata. The phyloseq object was filtered to remove OTUs observed <10 times, which removed more than half of observed otus (n=*83* OTUs removed and n=*81* kept, including n=*10* doubletons). The phyloseq object was further filtered to remove samples with very low sequence abundance (<200 sequences, n=*27* samples removed and n=*262* kept).

In *262* coral samples, we found XXXX sequences after quality filtering. *clade abundances here*


## Statistical Analysis 
Code will be avaible on git hub

# References {-}
<!-- Insert references  -->
```{r, warning=FALSE, message=FALSE, echo=FALSE} 
  write.bibtex(entry=NULL, file="knitcitations.bib", append=FALSE)
```
