---
bibliography: Paperpile.bib
csl: nature.csl
geometry: margin=2.5cm
linestretch: 2
mainfont: mathptmx
output:
  html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    theme: sandstone
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: no
    pandoc_args:
    - +RTS
    - -K64m
    - -RTS
  word_document:
    fig_caption: yes
    fig_width: 7.5
---

<!-- ## Install and source necessary packages-->
```{r, echo=FALSE, include=FALSE}


  # install.packages('knitcitations')
  # install.packages('bibtex')
  # install.packages('qdap')

  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries=FALSE)
  library(bibtex)
  
  # source('http://bioconductor.org/biocLite.R')
  # biocLite('phyloseq')
  # packageVersion('phyloseq')
  library(phyloseq)
  # library("qdap") # to be able to use word count
  library("plyr")
  require("dplyr")
  require("mosaic")
```
\pagebreak

TITLE {-}

Authors: Danielle C. Claar $^1$, Kristina L. Tietjen $^1$, Ruth D. Gates $^2$, Julia K. Baum $^1$

Institute: $^1$ Department of Biology, University of Victoria, PO BOX 1700 Station CSC, Victoria, British Columbia, V8W 2Y2, Canada
$^2$ Hawaii Institute of Marine Biology, 46-007 Lilipuna Road, Kaneohe, HI 96744, USA

Corresponding Author: Danielle C. Claar, Tel: (208) 250-0161, Email: dclaar@uvic.ca

Keywords: coral bleaching, heat stress, climate change, Symbiodinium

\pagebreak 

# Summary {-} 
Coral reefs, which already live on the edge of their thermal tolerance [@Sampayo2016-vd], are under acute threat from ocean warming [@Hughes2003-aj; @Hoegh-Guldberg2007-fh; @Baker2008-ky]. Corals live in symbiosis with an extraordinarily diverse genus of photosynthetic dinoflagellates (Symbiodinium spp.; [@Muscatine1977-pn; @Rowan1992-lg]). The symbiotic association and diversity of various taxa of *Symbiodinium* can be flexible over time [@Baker2003-ks; @Little2004-tm], and individual *Symbidoinium* taxa can range from parasites to mutualists in their interaction with their coral host [@Lesser2013-dj]. Warming causes the breakdown of coral symbiosis, causing coral "bleaching" when symbionts are expelled and the white coral skeleton is visible through the coral tissue [@Brown1997-mf]. Coral bleaching can lead to mortality, although corals can regain their symbionts after heat stress has abated [@Douglas2003-nr; @Stat2009-qq]. The 2015/16 El Niño is the worst pulse warming event on record in terms of severity and longevity [@Eakin:2016vf; @Heron2016-am], yet despite massive coral mortality, some corals show resilience to this extreme event (@Bauminprep). Here, we track coral symbioses and survival at the epicenter of this bleaching event (Kiritimati, Central Pacific), and show, contrary to our current paradigm of coral bleaching and recovery dynamics, that some corals have the capacity to re-establish symbiosis before heat stress subsides. Furthermore, we demonstrate potential mechanisms for coral survival and recovery, including the lack of preferential symbiont expulsion, and the effect of local human disturbance on  pre-bleaching symbiont community structure and the probability of coral survival. Together, these results show the potential for reef corals to survive extreme warming events, providing tentative hope for the survival of corals in the Anthropocene.

\pagebreak

# Main Text {-}
```{r, echo=FALSE, include=FALSE}
citep("Article")
```
Global coral bleaching is increasing, and the 2014-2017 event caused a catastrophic loss of corals around the globe. There was up to 95% mortality in some regions during the 1997/1998 El Niño event (Glynn 1993). The 2014-2017 global coral bleaching event caused coral bleaching across the world's oceans (Eakin 2016, Normile 2016), with up to 75% bleaching on some reefs in Hawaii, and at least some level of bleaching across 93% of the Great Barrier Reef (Minton et al 2015, GBRMPA 2016). The 2015-2016 El Niño, superimposed on nearly-ubiquitous tropical ocean warming, instigated the third global coral bleaching event (@Eakin:2016vf). 

The symbiosis between coral and their single-celled dinoflagellate symbionts, *Symbiodinium*, is the foundation of reef ecosystems, and a critical element of reef resilience [@Van_Oppen2006-qf]. The coral holobiont responds to environmental conditions, and is the unit that interacts with the broader reef community [@Gates2011-zy], supporting reef diversity and function across taxa *(or at a global scale?)*. There is much genetic, functional, and response diversity within the *Symbiodinium* genus. *Symbiodinium* types, considered putative species [@Pochon2010-jm], have distinct geographic distributions, host associations, and environmental optima [@Fabina2012-mm]. There are functional differences between *Symbiodinium* clades [@Stat2008-hk], and *Symbiodinium* associations can range from mutualistic to neutral to parasitic based on *Symbiodinium* type as well as environmental conditions [@Lesser2013-dj]. Recent advances in next-generation sequencing techniques have revealed cryptic genetic diversity within symbiotic *Symbiodinium* [@Quigley2014-zj; @Arif2014-kx; @Green2014-az], and has allowed for long-term genetic and ecological comparisons of symbiont community structure [@Edmunds_undated-fd].  
  
It is thought that corals may be able to survive thermal stress by changing their complement of symbionts to better suit environmental conditions. The adaptive bleaching hypothesis suggests that corals bleach in order to expel environmentally sub-optimal symbionts, followed by switching (picking up new symbionts from the environment) or shuffling (an internal change in dominant symbiont type or overall symbiont community structure) [@Buddemeier2004-se; @Buddemeier1993-sx; @Baker2001-vc; @Baker2003-ks]. There is evidence for both *Symbiodinium* shuffling (Rowan 2004) and switching [@Boulotte2016-dy]. However, what remains unclear is if and how frequently bleaching events can actually be considered adaptive. Changes in symbiotic function have been demonstrated due to shifts in the dominant *Symbiodinium* clade, but functional differences such as photosynthetic efficiency and bleaching resistance are also present among *Symbiodinium* types within a single clade [@Sampayo2008-tw; @Kemp2014-xj]. Mechanisms of thermal tolerance vary among coral taxa due to the variability in symbiotic flexibility between symbiotic generalists (coral species that associate with several *Symbiodinium* partners through space and/or time) and specialitsts (coral species that consistently associate with one or a limited number of *Symbiodinium* types). Clade D Symbiodinium are considered heat-tolerant symbionts [@Stat2010-zg]. Furthermore, repopulation of a coral host with clade D symbionts after a bleaching event is proposed to be a survival mechanism [@Berkelmans2006-rf; @Mieog2007-yy; @Silverstein2012-tm]. For example one study showed that a history of thermal stress increased the prevalence of clade D *Symbiodinium* in a generalist coral species, but did not instigate similar changes in two specialist coral species [@Stat2013-qp]. However, there is a tradeoff to housing Clade D Symbiodinium, as corals that house clade D symbionts may have slower growth rates [@Little2004-tm] or lower capacity for energy storage [@Jones2011-nf]. 
   
The current paradigm of coral bleaching and resilience is that as environmental stress (such as warming) increases, corals begin to lose their obligate symbionts (*Symbiodinium*) and "bleach" [@Gates1992-ew; @Douglas2003-nr]. Thermal stress is the primary cause for coral bleaching, and extreme or long-lasting warming causes a complete breakdown of the coral symbioses, leading to expulsion of all (or nearly all) *Symbiodinium* from the coral host tissue, leading to mortality [@Hoegh-Guldberg1999-rb]. Thermal stress can be exacerbated by other environmental stressors (Cooper et al 2011, Béraud et al 2013, Maina et al 2008). During bleaching, there is a window for recovery, that is, a certain amount of time during which the warming must cease and conditions must return to normal so that the coral can regain its symbionts. If the window for recovery passes without amelioration of environmental conditions, the coral will starve and die. (Cunning et al 2016, Putnam et al 2017). Survival through such an extreme heat event provides an exceptional opportunity to understand how some corals can withstand intense heat stress, and how corals in general might survive long-term warming. Remarkably, we find that some coral colonies were able to survive this prolonged heat stress by regaining their symbionts while temperatures were still elevated.
  
Here we show that despite unprecedented heat stress, some corals exhibited resilience and survived. Our study location, Kiritimati Atoll (Christmas Island, Kiribati, Central Equatorial Pacific, Coordinates: 2, -157.4), was at the epicenter of this extreme El Niño event. Thermal anomalies were severe on Kiritimati, rapidly exceeding NOAA Coral Reef Watch's Coral Bleaching Alert Level 1 and Alert Level 2 thresholds, reaching an unprecedented (@Hoegh-Guldberg2011-sl) 25.7 DHW over a year-long bleaching event, demolishing most of the reef (@Baum_inprep). Despite staggering losses caused by ocean warming, some corals have the capacity to be resilient to these increasingly frequent mass-bleaching events (Hughes et al 2017). Here, we assess coral symbiosis and survival during the massive 2015/2016 El Niño event. We tagged, sampled, and photographed the same coral colonies before, during, and immediately after the El Niño event. We assessed bleaching condition and survival for each coral colony, and used Illumina MiSeq ITS2 amplicon sequencing and 97% *de novo* OTU clustering to evaluate changes in *Symbiodinium* community structure. To investigate mechanisms underlying the ability of these corals to not only survive a year of continuous heat stress, but to recover in the interim, we assessed the relationship between human disturbance, pre-bleaching *Symbiodinium* community structure, and coral survival, as well as the timing of *Symbiodinium* community shifts throughout this El Niño event. 

We document, for the first time, corals that were able to visually recover from bleaching, and to regain their *Symbiodinium* communities during the course of an extreme heat stress event. These corals (family Faviidae; *Platygyra* sp. and *Favites* sp.) were bleached within two months of the onset of warming, but had visibly recovered after 10 consecutive months of intense warming (Fig. 1).

```{r, echo=FALSE, warning=FALSE, eval=TRUE,include=FALSE}
source("../figures/Figure1.R")
```
![Figure 1. Top: Heat stress, expressed as Degree Heating Weeks (DHW), on Kiritimati Island over the course of the 2015-2016 El Niño event. Corals are sensitive to temperatures warmer than 1&degC above their normal highest summertime mean sea surface temperature (SST), known as the bleaching threshold. DHW shows how much heat stress has accumulated in an area over the past twelve weeks by adding up any temperature exceeding the bleaching threshold during that time period. Horizontal lines show expected bleaching severity levels: 4 &degC (yellow line), NOAA Coral Reef Watch (CRW) Bleaching Alert Level 1, where significant coral bleaching is likely; 8 &degC (light orange line), NOAA CRW Bleaching Alert Level 2, where widespread bleaching and mortality may occur; 12 &degC (dark orange line), mass coral mortality (Hoegh-Guldberg 2011); 24 &degC (dark red line) "not experienced by reefs yet" (Hoegh-Guldberg 2011). Solid black line indicates *in situ* calculated DHW, and fill colors correspond to bleaching temperature thresholds. The dashed vertical gray line show the five sampling time points. Bottom: Photographs of a single tagged *Platygyra* coral colony (#99) at five sampling time points, showing the initially healthy colony bleached after two months of heat stress, 'recovered' to a normal brown colour after ten months of heat stress, and still alive six months post heat stress (Note: no photo is shown from the May 2015 sampling time point). NOTE: Add 'degree C-weeks' (which is the units for DHW to the y-axis of the plots; use degree symbol instead of the word degree)](../figures/Figure1.tiff)


Stochasticity in the rare *Symbiodinium* biosphere may build or weaken a coral's capacity for resilience. Corals commonly host background *Symbiodinium* types in low levels (Correa et al 2009), but sub-dominant *Symbiodinium* communities are often unstable (Coffroth et al 2010). In other systems, rare microbial species have been demonstrated to be disproportionally important to maintaining functional processes during environmental change (Shade et al 2014). The importance of rare *Symbiodinium* types is currently under debate, and these rare types may be commensal (symbionts that pass through coral's holobiont with no harm or gain for either partner), parasitic ("cheaters", or symbionts that take more than they give), or mutualistic (symbionts which support host function) (Parkinson et al 2015). Some research suggests that low-abundance *Symbiodinium* types have minimal functional significance to corals (Lee et al 2016), while other evidence supports the idea that the rare *Symbiodinium* biosphere is important for corals' response to climate change (Boulotte et al 2016) and that, shifts in *Symbiodinium* community diversity may have a larger influence on coral resilience than the evolution of symbiont thermal tolerance (Baskett et al 2010). We show that after two months of heat stress, fully-bleached corals retained approximately the same *Symbiodinium* community as they had before the bleaching event. This suggests that a wholesale breakdown of symbiosis occurred in bleached corals during this event, indicating a lack of preferential symbiont expulsion or exodus. Furthermore, we demonstrate that symbionts present in even very low abundances can play a critical role in coral survival and recovery. Some coral colonies recovered symbiosis with *Symbiodinium* types that were present in only a negligible amount before the bleaching event.
  
![Figure 2. *Symbiodinium* community composition at each of five time points, showing the shift in dominance from clade C to clade D over the course of the 2015-2016 El Niño event, for A. the entire pool of tagged *Platygyra* coral colonies at each of the five sampling time points (n= X - Y colonies per time point), B. the same individual tagged *Platygyra* coral colony (#99). -- This figure illustrates non-preferential explusion of clades c and d; and that symbiodinium that ar einitially extremely rare can play a critical role in coral resilience to heat stress.] 
  
Local protection is critical for coral symbiosis and survival. We show that corals living at different levels of local human disturbance had distinct symbiont communities that corresponded tightly to survivorship. This is in contrast to a recent study which concluded that particulate and dissolved nutrients do not reduce coral health at a colony scale (Rocker et al 2017). However, there is increasing evidence for local adaptation in corals (Howells et al 2012, Logan et al 2013, Dixon et al 2015). Our results suggest that some Kiritimati coral species may have the capacity to experience evolutionary rescue, defined as adaptation at a rate that allows an endangered population to survive the rate of environmental change (Orr & Unkless  2014, Carlson 2014). Our results suggest that the capacity for evolutionary rescue is tangibly related to local reef protection. Although massive bleaching events like this one will likely continue to cause catastrophic damage to coral reefs worldwide, mitigating local human disturbance can potentially help protect some coral species against a modest amount of ocean warming. 

![Figure 3. A. [Danielle to write this one: - will be the Constrained ordination plot showing groupings of *Symbiodinium* communities from individual *Platygyra* colonies, grouping into two distinct areas according to level of local disturbance....]; B. Bar plots showing *Symbiodinium* community composition for individual *Platygya* colonies at a single time point prior to the heat stress, from sites with high (top) and low (bottom) levels of local disturbance levels.] 
Figure 4. Potentially the symbiodinium network plot.
Figure 5. Potentially the rank abundance plot for Platy.....

# Methods {-}
*The Methods section should be written as concisely as possible but should contain all elements necessary to allow interpretation and replication of the results. As a guideline, Methods sections typically do not exceed 3,000 words. Detailed descriptions of methods already published should be avoided; a reference number can be provided to save space, with any new addition or variation stated. The Methods section should be subdivided by short bold headings referring to methods used and we encourage the inclusion of specific subsections for statistics, reagents and animal models. If further references are included in this section, the numbering should continue from the end of the last reference number in the rest of the paper and the list should accompany the additional Methods at the end of the paper. The Methods section cannot contain figures or tables (essential display items should be included in the Extended Data).*

## Field Sampling 

Kiritimati Atoll (Christmas Island), Kiribati is located in the Central Equatorial Pacific (1.9N 157W), at the center of the El Niño 3.4 region (a region which is used to quantify El Niño presence and strength). During the 2015/2016 El Niño event, Kiritimati experienced 10 months of sustained temperature stress, causing a mass bleaching and mortality event (@BauminPrep). 

## Temperature quantification
Temperature loggers (Sea-Bird 56) were deployed around the island at 10-12m depth from 2011-2016 to measure *in situ* thermal stress.

##Coral Tagging and sampling  
In August/September 2014 colonies of *Platygyra* sp. and *Favites pentagona* were tagged along a 60m transect at 10-12m depth at 15 different sites around the atoll.  A photo was taken of each coral to record colony measurments and characteristics (i.e. bleaching). The tagged coral colonies were resampled twice more before (January/February 2015, April/May 2015), once during (July 2015), and once near the end (March 2016) of the El Niño warming. Some tagged coral colonies were lost due to storm damage, and new coral colonies were tagged to replenish the total number of surveyed colonies. Not all sites were visited during all field seasons, and some site surveys were only partially completed during some field seasons due to inclement weather conditions.

Corals were sampled by... processed like this... Coral tissue samples were preserved in Guanidinium buffer (50% w/v guanidinium isothiocyanate; 50 mM Tris pH 7.6; 10 µM EDTA; 4.2% w/v sarkosyl; 2.1% v/v β-mercaptoethanol) and stored at 4 *degrees* until extraction.

## Pre-processing and sequencing
DNA extraction was performed using a guanidinium-based extraction protocol [@Stat2009-qq; Cunning2017-sc; Cunning2015-mt] with the modification that the DNA pellet was washed with 70% ethanol three times rather than once.

Library Prep- Amy's method of library prep, include cleanup, Illumina Sequencing information (barcodes, etc) for sequencing of ITS2 amplicons (‘itsD’ and ‘its2rev2’ primers from [@Stat2009-qq] 
Illumina MiSeq platform with 2×300 paired-end read chemistry.
HIMB - Amy

ITS2 region - it's annoying, but it's the best we've got right now smith et al 2017

A total of 289 samples were prepared for sequencing, and *XXXX* of these samples were successfully amplified, sequenced, and used in downstream analyses.



## Bioinformatics
We conducted quality filtering of raw reads (in .fastq format) first using iu-filter-quality-bokulich implemented in Illumina-Utils [@Bokulich2013-cm; Eren2013-yg], followed by paired-end sequence merging via iu-merge-pairs (also in Illumina-Utils, [Eren2013-yg]), with a maximum mismatch of three bases between the forward and reverse reads. After quality filtering, sequence processing and identification was performed following all specifications of [@Cunning2017-sc]; chimeric sequences were removed, primers were trimmed, sequences from each sample were clustered independently at 97% similarity using UCLUST [@Edgar2010-zl] implemented in QIIME [@Caporaso2010-yl] and resulting OTUs were collapsed at 100% identity across samples, sequences were aligned using the Needleman-Wunsch global alignment algorithm (Biostrings package, [@Pages2017-ie]) in R [@R_Development_Core_Team2008-sp], and sequences were named using a reference database. 

The Phyloseq package [@McMurdie2013-hf] in R was used to store and analyze OTU tables, taxonomic information, and sample metadata. The phyloseq object was filtered to remove OTUs observed <10 times, which removed more than half of observed otus (n=*83* OTUs removed and n=*81* kept, including n=*10* doubletons). The phyloseq object was further filtered to remove samples with very low sequence abundance (<200 sequences, n=*27* samples removed and n=*262* kept).

In *262* coral samples, we found XXXX sequences after quality filtering. *clade abundances here*


## Statistical Analysis 
Code will be avaible on git hub

# References {-}
<!-- Insert references  -->
```{r, warning=FALSE, message=FALSE, echo=FALSE} 
  write.bibtex(entry=NULL, file="knitcitations.bib", append=FALSE)
```
