---
title: "KI Platygyra Manuscript"
author: "Danielle C Claar, Kristina L Tietjen, Ruth D Gates, Julia K Baum"
# date: 'r format(Sys.time(), "%d %B, %Y")'
output:
  html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    theme: cerulean
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
csl: nature.csl
bibliography: PapersLibrary_Bibtex.bib
---

<!-- ## Install and source necessary packages-->
```{r, echo=FALSE, include=FALSE}


  # install.packages('knitcitations')
  # install.packages('bibtex')
  # install.packages('qdap')

  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries=FALSE)
  library(bibtex)
  
  # source('http://bioconductor.org/biocLite.R')
  # biocLite('phyloseq')
  # packageVersion('phyloseq')
  library(phyloseq)
  # library("qdap") # to be able to use word count
  library("plyr")
  require("dplyr")
  require("mosaic")
```

# Summary
*Let's leave this for now and come back to it once we have rest of manuscript honed* Coral reefs, which already live on the edge of their thermal tolerance (@Sampayo2016-vd), are under acute threat from ocean warming (@Hughes2003-aj, @Hoegh-Guldberg2007-fh, @Baker2008-ky). Corals live in symbiosis with an extraordinarily diverse genus of photosynthetic dinoflagellates (Symbiodinium spp.; @Muscatine1977-pn, @Rowan1992-lg). The symbiotic association and diversity of various taxa of *Symbiodinium* can be flexible over time (@Baker2003-ks, @Little2004-tm, but see @Thornhill2006-cs, @Thornhill2009-jh), and individual *Symbidoinium* taxa can range from parasites to mutualists in their interaction with their coral host (@Lesser2013-dj). Warming causes the breakdown of coral symbiosis, causing coral "bleaching" when symbionts are expelled and the white coral skeleton is visible through the coral tissue (@Brown1997-mf). Coral bleaching can lead to mortality, although corals can regain their symbionts after heat stress has abated (@Douglas2003-nr, @Stat2009-qq). The 2015/16 El Ni?o is the worst pulse warming event on record in terms of severity and longevity (@Eakin:2016vf, @Heron2016-am), yet despite massive coral mortality, some corals show resilience to this extreme event (@Bauminprep). Here, we track coral symbioses and survival at the epicenter of this bleaching event (Kiritimati, Central Pacific), and show, contrary to our current paradigm of coral bleaching and recovery dynamics, that some corals have the capacity to re-establish symbiosis before heat stress subsides. Furthermore, we demonstrate potential mechanisms for coral survival and recovery, including the lack of preferential symbiont expulsion, and the effect of local human disturbance on  pre-bleaching symbiont community structure and the probability of coral survival. Together, these results show the potential for reef corals to survive extreme warming events, providing tentative hope for the survival of corals in the Anthropocene.

# Main Text
```{r, echo=FALSE, include=FALSE}
citep("Article")
```
*JKB, 29 June 2017 - Before we proceed further, let's bullet point, our main findings as well as the prevailing views from the field. i.e. We need to very clearly lay out for a highly educated general audience what the expectation was for how corals would respond to heat stress and how what we found differs from the expectation. The expectation part needs to be greatly strengthened in the current draft i.e. we need to set up 'the mystery'*

*Here's what I see as the main discoveries:* 

1. Despite unprecedented heat stress, we show that some corals exhibited resilience and survived. Survival through such an extreme heat event provides an exceptional opportunity to understand how some corals can withstand intense heat stress, and how corals in general might survive long-term warming. Remarkably, we find that some coral colonies were able to survive this prolonged heat stress by regaining their symbionts while temperatures were still elevated.

2. After two months of heat stress, fully-bleached corals retained approximately the same *Symbiodinium* community as they had before the bleaching event. This suggests that a wholesale breakdown of symbiosis occurred in bleached corals during this event, indicating a lack of preferential symbiont expulsion or exodus.

3. Symbionts present in even very low abundances can play a critical role in coral survival and recovery. Some coral colonies recovered symbiosis with *Symbiodinium* types that were present in only a negligible amount before the bleaching event.

4. Local protection is critical for coral symbiosis and survival. Corals living at different levels of local human disturbance had distinct symbiont communities that corresponded tightly to survivorship. 

*Danielle to write out here a section of bullet points on - 'What is expected to happen to corals during heat stress?' (this can include bullets on prevailing viewpoints, bullets on observations from previous heat stress events or from this heat stress event in other locations (e.g. Hughes - GBR), along with the relevant citations).*

For example:
1. The prevailing viewpoint is that corals subject to elevated temperatures expel their symbionts and bleach.... (general statement here, including over what time period, and that they can recover their symbionts once the heat stress subsides). General points about the intensity / length /DHW that corals are thought to bleach at, are thought to survive through, or conversely thought to die from e.g. 'Corals do not typically survive more than X months of Y-Z degrees of elevated temperatures' - do we know what the most intense heat stress is that corals have previously been documented to survive through?

2. Prevailing viewpiont is that when coral colonies expel their symbionts, they preferentialy expel the sub-optimal ones. This typically means expelling symbions in Clade C. This has been shown in (explain studies....)

3. Prevailing viewpoint is that very rare symbionts are unstable and don't contribute much to coral colony fitness..... thus if corals are to survive heat stress it is generally believed that they either need to start with heat-tolerant D's being dominant (?, or at least quite common) or would need to acquire them from the surrounding environment. 
-Please confirm that we have the data from your water samples to show that those D's were not just floating around in the water column, but rather the only place they existed was in very very low #s within the colonies themselves?  

4. What is known / expected about local protection's influence on symbiont community? influence on survival through heat stress?

Danielle's Notes:  
  
1.	*Coral symbiosis is the foundation of coral reef ecosystems* (van Oppen & Gates 2006, Muller-Parker 2015).  
  
2.	*The coral holobiont responds to environmental conditions, and is the unit that interacts with the broader reef community* (Gates and Ainsworth 2011)  
  
3.	*There is much genetic, functional, and response diversity within Symbiodinium*  
Although Symbiodinium is a single genus, it contains diversity similar to diversity found within other dinoflagellate orders (Rowan and Powers 1992), and is divided into nine clades and hundreds of types (Rodrigues-Lanetty et al 2011, van Oppen et al 2005, Franklin et al 2012, Lesser et al 2013, Tonk et al 2013). There is much debate about species classifications within Symbiodinium, which has hampered species naming within Symbiodinium (Stat et al 2012, Smith 2017 but see LaJeunesse 2001). Symbiodinium types have distinct geographic distributions, host associations, and environmental optima (Fabina et al 2012). Furthermore, total and relative abundance of Symbiodinium can vary among coral colonies, across environmental gradients, and over time (Fitt 2000), with increased Symbiodinium abundance leading to increased environmental sensitivity and bleaching risk (Cunning and Baker 2013). There are functional differences between Symbiodinium clades (Stat et al 2008), and Symbiodinium associations can range from mutualistic to neutral to parasitic based on Symbiodinium type as well as environmental conditions (Lesser et al 2013). Next-generation sequencing has revealed cryptic genetic diversity within symbiotic Symbiodinium (Kenkel et al 2013, Nelson et al 2014, Quigley et al 2014), and has allowed for long-term genetic and ecological comparisons of symbiont community structure (Edmunds et al 2014).  
  
4.	*Corals exhibit varying levels of symbiotic flexibility, but this flexibility comes with functional tradeoffs.*  
Corals that host flexible symbioses (generalists) may be more sensitive to environmental perturbations than those with intimate symbioses (specialists) (Putnam et al 2012). A history of thermal stress increased the prevalence of clade D Symbiodinium in a generalist coral species, but did not instigate similar changes in two specialist coral species (Stat et al 2013). Although the prevalence of clade D Symbiodinium increases during thermal stress and may increase thermal tolerance (Berkelmans and van Oppen 2006), corals that house clade D symbionts may have slower growth rates (Little et al 2004) or lower energy storage (Jones and Berkelmans 2011).  
  
5.	*The current paradigm of coral bleaching and recovery states that the stress must cease for coral to regain their symbiosis*  
Coral bleaching is the loss of obligate symbionts (Symbiodinium) from the coral tissue (Gates et al 1992, Douglas 2003). Thermal stress is the primary cause for coral bleaching, and can cause not only the breakdown of coral symbioses, but also cause coral mortality (Hoegh-Guldberg 1999, Abrego et al 2012, Stat et al 2013, Baker 2013). Thermal stress can be exacerbated by other environmental stressors (Cooper et al 2011, Béraud et al 2013, Maina et al 2008), and in turn, exacerbates ocean acidification (Gibbin et al 2015). The current paradigm of coral bleaching and resilience is that as environmental stress  (such as warming) increases, corals begin to bleach. Extreme or long-lasting warming causes a complete breakdown of the coral symbioses, leading to expulsion of all (or nearly all) Symbiodinium from the coral host tissue. It has been shown that during bleaching, there is a window for recovery, that is, a certain amount of time during which the warming must cease and conditions must return to normal so that the coral can regain its symbionts. If the window for recovery passes without amelioration of the environmental conditions, the coral will starve and die. (Cunning et al 2016, Putnam et al 2017).  
  
6.	*The adaptive bleaching hypothesis provides a testable hypothesis for bleaching causes and consequences, but hasn't found extensive support*  
The adaptive bleaching hypothesis suggests that corals bleach in order to expel environmentally sub-optimal symbionts, followed by switching (picking up new symbionts from the environment) or shuffling (an internal change in dominant symbiont type or overall symbiont community structure) (Buddemeier and Fautin 1993, Baker 2001, Baker 2003, Buddemeier et al 2004).  
  
7.	*The "transient microbiome" assembled by environmental anomalies can undergo rapid changes (Putnam et al 2017), providing symbiotic stochasticity which may build or weaken a coral's capacity for resilience.*  
The "transient microbiome" assembled by environmental anomalies can undergo rapid changes (Putnam et al 2017), providing symbiotic stochasticity which may strengthen or weaken a coral's capacity for resilience. While corals have been shown to change symbiotic partners during a bleaching event, there is often a quick return to pre-bleaching Symbiodinium communities after recovery (Thornhill 2005, LaJeunesse et al 2010), and persistence of stress-related changes to Symbiodinium community structure may require sustained environmental pressure (Baird et al 2007). Corals commonly host background Symbiodinium types in low levels (Correa et al 2009), but sub-dominant Symbiodinium communities are often unstable (Coffroth et al 2010). However, shifts in Symbiodinium community diversity may still have a larger influence on coral resilience than the evolution of symbiont thermal tolerance (Baskett et al 2010). 
  
8.	*Global coral bleaching is increasing, and the 2014-2017 event caused a catastrophic loss of corals around the globe.*  
There was up to 95% mortality in some regions during the 1997/1998 El Niño event (Glynn 1993). The 2014-2017 global coral bleaching event caused coral bleaching across the world's oceans (Eakin 2016, Normile 2016), with up to 75% bleaching on some reefs in hawaii, and at least some level of bleaching across 93% of the Great Barrier Reef (Minton et al 2015, GBRMPA 2016). Despite these staggering losses, some corals have the capacity to be resilient to these increasingly frequent mass-bleaching events (Hughes et al 2017).   
  
9.	*Our results suggest that some Kiritimati coral species may experience evolutionary rescue*   
Our results suggest that some Kiritimati coral species may experience evolutionary rescue, defined as adaptation at a rate that allows an endangered population to survive the rate of environmental change (Orr & Unkless  2014, Carlson 2014).   
  
10.	*There is evidence for local adaptation in corals* (Howells et al 2012, Logan et al 2013, Dixon et al 2015).  
  
11.	*Thick-tissue corals may be less susceptible to coral bleaching and mortality due to their tissue biomass and associated energy reserves.*  
Thick-tissue corals (such as Platygyra) may be less susceptible to coral bleaching and mortality due to their tissue biomass and associated energy reserves (Loya et al 2001). Furthermore, self-shading can protect "understory" Symbiodinium, those which reside deeper in the coral's tissue and consequently experiences less light stress, which can provide a symbiont reserve for repopulation and recovery (Kemp et al 2014).  
  


*Main text draft*
The symbiosis between coral and their single-celled dinoflagellate symbionts, *Symbiodinium*, is the foundation of reef ecosystems, and a critical element of reef resilience (@Van_Oppen2006-qf). Corals host a diverse community of *Symbiodinium*, ranging along a continuum from 'selfish opportunistic symbionts' (e.g. some clade D *Symbiodinium*) which are better suited to sustained environmental stress, to 'intimately evolved symbionts' which provide exceptional amounts of nutrition to their coral host (@Stat:2006ww , @Lesser:2013tu). Diversity with the genus *Symbiodinium* is high, comparable to genetic variability among orders in other dinoflagellate taxa (@Rowan1992-lg), and is divided hierarchically into clades, subclades, and types. Symbiotic flexibility and stability can be variable both within and among coral species (@Putnam2012-om, CITE?), and this can be a driver in determining "winers and losers" (@Loya2001-gb) during coral bleaching events. Coral bleaching is the breakdown of symbiosis, where *Symbiodinium* are expelled en masse from the tissues of their coral host. A coral's susceptibility and resilience to bleaching is, in part, determined by fine-scale variability in their compliment of associated symbionts (@Sampayo2008-tw). Corals have been observed to recover from bleaching only if the underlying stress, such as ocean warming, abates.

Ocean warming events can cause massive losses of coral cover (CITE,CITE). The 2015-2016 El Ni?o, superimposed on nearly-ubiquitous tropical ocean warming, instigated the third global coral bleaching event (@Eakin:2016vf). Our study location, Kiritimati Atoll (Christmas Island, Kiribati, Central Equatorial Pacific, Coordinates: 2, -157.4), was at the epicenter of this extreme El Ni?o event. Thermal anomalies were severe on Kiritimati, rapidly exceeding NOAA Coral Reef Watch's Coral Bleaching Alert Level 1 (4 Degree Heating Weeks, DHW, a metric of cumulative thermal stress) and Alert Level 2 (8 DHW) thresholds, reaching an unprecedented (@Hoegh-Guldberg2011-sl) 25.7 DHW over a year-long bleaching event, demolishing most of the reef (@Baum_inprep). Despite the massive mortality resulting from this extreme heat stress event, some corals survived.

Here, we assess coral symbiosis and survival during the massive 2015/2016 El Ni?o event. We tagged, sampled, and photographed the same coral colonies before, during, and immediately after the El Ni?o event. We assessed bleaching condition and survival for each coral colony, and used Illumina MiSeq ITS2 amplicon sequencing and 97% *de novo* OTU clustering to evaluate changes in *Symbiodinium* community structure. To investigate mechanisms underlying the ability of these corals to not only survive a year of continuous heat stress, but to recover in the interim, we assessed the relationship between human disturbance, pre-bleaching *Symbiodinium* community structure, and coral survival, as well as the timing of *Symbiodinium* community shifts throughout this El Ni?o event. We document, for the first time, corals that were able to visually recover from bleaching, and to regain their *Symbiodinium* communities during the course of an extreme heat stress event. These corals (family Faviidae; *Platygyra* sp. and *Favites* sp.) were bleached within two months of the onset of warming, but had visibly recovered after 10 consecutive months of intense warming. Previously, corals have been shown to recover from bleaching only after the external stress (e.g. warming) has subsided (CITE), implying that longer and more frequent stressors spell disaster for reefs worldwide. This unprecedented resilience mechanism...

```{r, echo=FALSE, warning=FALSE, eval=TRUE,include=FALSE}
source("../figures/Figure1.R")
```
![Figure 1](../figures/Figure1.tiff)

*Symbiodinium* clade predicts survival, and local human disturbance predisposes coral to have "suboptimal" clades.
Opposed to other study which showed that only those with D survived, C didn't, and that those which started with undetectable D and switched to C still died. Bay et al 2016 (threshold densities study)
For *Platygyra*, Corals with clade C survived, and corals with clade D died. *Symbiodinium* community structure was stable within coral colonies before the onset of bleaching. The pre-bleaching distribution of *Symbiodinium* clade was associated with local human disturbance level, with corals at high-impact sites housing clade D as the dominant symbiont, and those at low-impact sites housing clade C as the dominant symbiont. 

We do know that corals house background symbionts in low abundances (@Correa:2009jy, all the recent ngs studies...), but these relationships have been described as unstable (@Coffroth:2010ju, more cites?). Silverstein et al 2012 Specificity is rarely absolute
Despite the consensus that coral symbioses are integral to coral resistance to and resilience from bleaching, it is thought that most background *Symbiodinium* types have minimal functional significance for corals (@Lee2016-ss), and . .
 We demonstrate that tiny abundances are important - many corals who switched from C to D during the event had extremely low or undetectable sequence counts of clade D *Symbiodinium* before the bleaching event, but their recovery symbiont communities were dominated by clade D *Symbiodinium* 

Our study also provides insight into the timing of symbiont community shifts in the field. *Symbiodinium* communities remained largely consistent within an individual coral colony from before-bleaching time points until many corals were mostly or entirely bleached. Significant symbiont community change only occurred after the coral was no longer visibly bleached, suggesting that in this bleaching event, corals experienced a wholesale expulsion of all symbionts (with no preferential expulsion), followed by competitive repopulation of the coral host. This adds evidence to support the theory that clade D *Symbiodinium* are competitively dominant when symbiont density is low in warm conditions (cite people who have said this before).
We used to think that bleaching might be good - ABH says that corals bleach in order to expel suboptimal *Symbiodinium* types in exchange for optimal symbionts during the new conditions (@Buddemeier:1993bb, Baker:2001bf, @Buddemeier:2004vj). *Switching and shuffling (@Baker:2003uo)* *And we do know that some symbio are "better" than others* *And then we said that bleaching is definitely bad* *But at least we do know that it allows changes to occur in the *Symbiodinium* community structure* 

Although massive bleaching events like this one will likely continue to cause catastrophic damage to coral reefs worldwide, mitigating local human disturbance can potentially help protect some coral species against a modest amount of ocean warming. Elucidating the mechanisms underlying changes in coral-symbiont interactions is essential to understanding the ability of the coral symbiome to adapt to the multiple stressors they now face. *CITE Shifting paradigms in restoration of the world's coral reefs @Van_Oppen2017-vv*

# Methods
*The Methods section should be written as concisely as possible but should contain all elements necessary to allow interpretation and replication of the results. As a guideline, Methods sections typically do not exceed 3,000 words. Detailed descriptions of methods already published should be avoided; a reference number can be provided to save space, with any new addition or variation stated. The Methods section should be subdivided by short bold headings referring to methods used and we encourage the inclusion of specific subsections for statistics, reagents and animal models. If further references are included in this section, the numbering should continue from the end of the last reference number in the rest of the paper and the list should accompany the additional Methods at the end of the paper. The Methods section cannot contain figures or tables (essential display items should be included in the Extended Data).*

## Field Information

Kiritimati Atoll, Kiribati is located in the Central Equatorial Pacific (1.9N 157W), at the center of the El Ni?o 3.4 region (used to quantify El Ni?o presence and strength). During the 2015/2016 El Ni?o event, Kiritimati experienced 10 months of sustained temperature stress, causing a mass bleaching and mortality event (@BauminPrep). 

Temperature loggers (Sea-Bird 56) were deployed around the island at 10-12m depth from 2011-2016 to measure *in situ* thermal stress.

##Coral Tagging and sampling  
In August/September 2014 seven taxa of coral were tagged along a 60m transect at 10-12m depth at 10 different sites around the atoll.  At each site, 12 colonies of each focal species (*Porites lobata*, *Montipora aequituberculata*, and *Pocillopora eyedouxi*) were tagged and a total of 18 colonies of a varying assortment of *Platygyra*, *Favities*, *Favia*, and *Hydnophora microconos* were tagged. A photo of each coral with a ruler beside it was taken for colony measurments and characterists (i.e. bleaching) to be analyzed in the lab. The tissue samples were stored in a cooler on the boat and at night they were rinsed with fresh water and alliquoted into centrifuge tubes with Guanidinium, then stored at 4 (I havent looked to see how to put in a degree symbol yet) C.  The tagged coral colonies were resampled once more before the El Ni?o (January/February 2015), twice during (May/April 2015 and July 2015), and once near the end (March 2016). Due to either storm waves breaking and moving coral or the tags dissapearing some of the corals were unable to be located and new corals were tagged to replenish numbers **I(KT) dont like this sentence - still working on how to word it*.


## Pre-processing and sequencing
DNA extraction was performed as in Stat et al. (2009) with a few modifications. The DNA pellet was washed with 70% ethanol three times, and after the final wash was removed, the remaining ethanol was allowed to evaporate off before the DNA pellet was resuspended in Tris buffer (0.1M pH8).
Cleanup by Amy here
Library Prep
sequencing exact information


ITS2 region - it's annoying, but it's the best we've got right now
PCR and Cleanup - Amy's method of PCR and cleanup
Library Prep - Amy's method of library prep, include Illumina Sequencing information (barcodes, etc)

## Bioinformatics
Sequence QC - boku, then merge with illumina utils, max mismatch=3
Sequence clustering - denovo clustering using UCLUST in QIIME, then compare to reference database to assign taxonomy


## Statistical Analysis 
Alpha diversity of sequence reads, co-occurence?, beta diversity?
Code will be avaible on git hub

# References
<!-- Insert references  -->
```{r, warning=FALSE, message=FALSE, echo=FALSE} 
  write.bibtex(entry=NULL, file="knitcitations.bib", append=FALSE)
```
