---
title: "KI Platygyra Manuscript"
author: "Danielle C Claar, Kristina L Tietjen, Ruth D Gates, Julia K Baum"
# date: 'r format(Sys.time(), "%d %B, %Y")'
output:
  html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    theme: cerulean
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
csl: nature.csl
bibliography: PapersLibrary_Bibtex.bib
---

<!-- ## Install and source necessary packages-->
```{r, echo=FALSE, include=FALSE}


  # install.packages('knitcitations')
  # install.packages('bibtex')
  # install.packages('qdap')

  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries=FALSE)
  library(bibtex)
  
  # source('http://bioconductor.org/biocLite.R')
  # biocLite('phyloseq')
  # packageVersion('phyloseq')
  library(phyloseq)
  # library("qdap") # to be able to use word count
  library("plyr")
  require("dplyr")
  require("mosaic")
```

# Summary
Coral reefs, which already live on the edge of their thermal tolerance (@Sampayo2016-vd), are under acute threat from ocean warming (@Hughes2003-aj, @Hoegh-Guldberg2007-fh, @Baker2008-ky). Corals live in symbiosis with an extraordinarily diverse genus of photosynthetic dinoflagellates (Symbiodinium spp.; @Muscatine1977-pn, @Rowan1992-lg). The symbiotic association and diversity of various taxa of *Symbiodinium* can be flexible over time (@Baker2003-ks, @Little2004-tm, but see @Thornhill2006-cs, @Thornhill2009-jh), and individual *Symbidoinium* taxa can range from parasites to mutualists in their interaction with their coral host (@Lesser2013-dj). Warming causes the breakdown of coral symbiosis, causing coral "bleaching" when symbionts are expelled and the white coral skeleton is visible through the coral tissue (@Brown1997-mf). Coral bleaching can lead to mortality, although corals can regain their symbionts after heat stress has abated (@Douglas2003-nr, @Stat2009-qq). The 2015/16 El Niño is the worst pulse warming event on record in terms of severity and longevity (@Eakin:2016vf, @Heron2016-am), yet despite massive coral mortality, some corals show resilience to this extreme event (@Bauminprep). Here, we track coral symbioses and survival at the epicenter of this bleaching event (Kiritimati, Central Pacific), and show, contrary to our current paradigm of coral bleaching and recovery dynamics, that some corals have the capacity to re-establish symbiosis before heat stress subsides. Furthermore, we demonstrate potential mechanisms for coral survival and recovery, including the lack of preferential symbiont expulsion, and the effect of local human disturbance on  pre-bleaching symbiont community structure and the probability of coral survival. Together, these results show the potential for reef corals to survive extreme warming events, providing tentative hope for the survival of corals in the Anthropocene.

# Main Text
```{r, echo=FALSE, include=FALSE}
citep("Article")
```

The symbiosis between coral and their single-celled dinoflagellate symbionts, *Symbiodinium*, is the foundation of reef ecosystems, and a critical element of reef resilience (@Van_Oppen2006-qf). Corals host a diverse community of *Symbiodinium*, ranging along a continuum from 'selfish opportunistic symbionts' (e.g. some clade D *Symbiodinium*) which are better suited to sustained environmental stress, to 'intimately evolved symbionts' which provide exceptional amounts of nutrition to their coral host (@Stat:2006ww , @Lesser:2013tu). Diversity with the genus *Symbiodinium* is high, comparable to genetic variability among orders in other dinoflagellate taxa (@Rowan1992-lg), and is divided hierarchically into clades, subclades, and types. Symbiotic flexibility and stability can be variable both within and among coral species (@Putnam2012-om, CITE?), and this can be a driver in determining "winers and losers" (@Loya2001-gb) during coral bleaching events. Coral bleaching is the breakdown of symbiosis, where *Symbiodinium* are expelled en masse from the tissues of their coral host. A coral's susceptibility and resilience to bleaching is, in part, determined by fine-scale variability in their compliment of associated symbionts (@Sampayo2008-tw). Corals have been observed to recover from bleaching only if the underlying stress, such as ocean warming, abates.

Ocean warming events can cause massive losses of coral cover (CITE,CITE). The 2015-2016 El Niño, superimposed on nearly-ubiquitous tropical ocean warming, instigated the third global coral bleaching event (@Eakin:2016vf). Our study location, Kiritimati Atoll (Christmas Island, Kiribati, Central Equatorial Pacific, Coordinates: 2, -157.4), was at the epicenter of this extreme El Niño event. Thermal anomalies were severe on Kiritimati, rapidly exceeding NOAA Coral Reef Watch's Coral Bleaching Alert Level 1 (4 Degree Heating Weeks, DHW, a metric of cumulative thermal stress) and Alert Level 2 (8 DHW) thresholds, reaching an unprecedented (@Hoegh-Guldberg2011-sl) 25.7 DHW over a year-long bleaching event, demolishing most of the reef (@Baum_inprep). Despite the massive mortality resulting from this extreme heat stress event, some corals survived.

Here, we assess coral symbiosis and survival during the massive 2015/2016 El Niño event. We tagged, sampled, and photographed the same coral colonies before, during, and immediately after the El Niño event. We assessed bleaching condition and survival for each coral colony, and used Illumina MiSeq ITS2 amplicon sequencing and 97% *de novo* OTU clustering to evaluate changes in *Symbiodinium* community structure. To investigate mechanisms underlying the ability of these corals to not only survive a year of continuous heat stress, but to recover in the interim, we assessed the relationship between human disturbance, pre-bleaching *Symbiodinium* community structure, and coral survival, as well as the timing of *Symbiodinium* community shifts throughout this El Niño event. We document, for the first time, corals that were able to visually recover from bleaching, and to regain their *Symbiodinium* communities during the course of an extreme heat stress event. These corals (family Faviidae; *Platygyra* sp. and *Favites* sp.) were bleached within two months of the onset of warming, but had visibly recovered after 10 consecutive months of intense warming. Previously, corals have been shown to recover from bleaching only after the external stress (e.g. warming) has subsided (CITE), implying that longer and more frequent stressors spell disaster for reefs worldwide. This unprecedented resilience mechanism...

```{r, echo=FALSE, warning=FALSE, eval=TRUE,include=FALSE}
source("../figures/Figure1.R")
```
![Figure 1](../figures/Figure1.tiff)

*Symbiodinium* clade predicts survival, and local human disturbance predisposes coral to have "suboptimal" clades.
Opposed to other study which showed that only those with D survived, C didn't, and that those which started with undetectable D and switched to C still died. Bay et al 2016 (threshold densities study)
For *Platygyra*, Corals with clade C survived, and corals with clade D died. *Symbiodinium* community structure was stable within coral colonies before the onset of bleaching. The pre-bleaching distribution of *Symbiodinium* clade was associated with local human disturbance level, with corals at high-impact sites housing clade D as the dominant symbiont, and those at low-impact sites housing clade C as the dominant symbiont. 

We do know that corals house background symbionts in low abundances (@Correa:2009jy, all the recent ngs studies...), but these relationships have been described as unstable (@Coffroth:2010ju, more cites?). Silverstein et al 2012 Specificity is rarely absolute
Despite the consensus that coral symbioses are integral to coral resistance to and resilience from bleaching, it is thought that most background *Symbiodinium* types have minimal functional significance for corals (@Lee2016-ss), and . .
 We demonstrate that tiny abundances are important - many corals who switched from C to D during the event had extremely low or undetectable sequence counts of clade D *Symbiodinium* before the bleaching event, but their recovery symbiont communities were dominated by clade D *Symbiodinium* 

Our study also provides insight into the timing of symbiont community shifts in the field. *Symbiodinium* communities remained largely consistent within an individual coral colony from before-bleaching time points until many corals were mostly or entirely bleached. Significant symbiont community change only occurred after the coral was no longer visibly bleached, suggesting that in this bleaching event, corals experienced a wholesale expulsion of all symbionts (with no preferential expulsion), followed by competitive repopulation of the coral host. This adds evidence to support the theory that clade D *Symbiodinium* are competitively dominant when symbiont density is low in warm conditions (cite people who have said this before).
We used to think that bleaching might be good - ABH says that corals bleach in order to expel suboptimal *Symbiodinium* types in exchange for optimal symbionts during the new conditions (@Buddemeier:1993bb, Baker:2001bf, @Buddemeier:2004vj). *Switching and shuffling (@Baker:2003uo)* *And we do know that some symbio are "better" than others* *And then we said that bleaching is definitely bad* *But at least we do know that it allows changes to occur in the *Symbiodinium* community structure* 

Although massive bleaching events like this one will likely continue to cause catastrophic damage to coral reefs worldwide, mitigating local human disturbance can potentially help protect some coral species against a modest amount of ocean warming. Elucidating the mechanisms underlying changes in coral-symbiont interactions is essential to understanding the ability of the coral symbiome to adapt to the multiple stressors they now face. *CITE Shifting paradigms in restoration of the world's coral reefs @Van_Oppen2017-vv*

# Methods
*The Methods section should be written as concisely as possible but should contain all elements necessary to allow interpretation and replication of the results. As a guideline, Methods sections typically do not exceed 3,000 words. Detailed descriptions of methods already published should be avoided; a reference number can be provided to save space, with any new addition or variation stated. The Methods section should be subdivided by short bold headings referring to methods used and we encourage the inclusion of specific subsections for statistics, reagents and animal models. If further references are included in this section, the numbering should continue from the end of the last reference number in the rest of the paper and the list should accompany the additional Methods at the end of the paper. The Methods section cannot contain figures or tables (essential display items should be included in the Extended Data).*

## Field Information

Kiritimati Atoll, Kiribati is located in the Central Equatorial Pacific (1.9N 157W), at the center of the El Niño 3.4 region (used to quantify El Niño presence and strength). During the 2015/2016 El Niño event, Kiritimati experienced 10 months of sustained temperature stress, causing a mass bleaching and mortality event (@BauminPrep). 

Temperature loggers (Sea-Bird 56) were deployed around the island at 10-12m depth from 2011-2016 to measure *in situ* thermal stress.

##Coral Tagging and sampling  
In August/September 2014 seven taxa of coral were tagged along a 60m transect at 10-12m depth at 10 different sites around the atoll.  At each site, 12 colonies of each focal species (*Porites lobata*, *Montipora aequituberculata*, and *Pocillopora eyedouxi*) were tagged and a total of 18 colonies of a varying assortment of *Platygyra*, *Favities*, *Favia*, and *Hydnophora microconos* were tagged. A photo of each coral with a ruler beside it was taken for colony measurments and characterists (i.e. bleaching) to be analyzed in the lab. The tissue samples were stored in a cooler on the boat and at night they were rinsed with fresh water and alliquoted into centrifuge tubes with Guanidinium, then stored at 4 (I havent looked to see how to put in a degree symbol yet) C.  The tagged coral colonies were resampled once more before the El Niño (January/February 2015), twice during (May/April 2015 and July 2015), and once near the end (March 2016). Due to either storm waves breaking and moving coral or the tags dissapearing some of the corals were unable to be located and new corals were tagged to replenish numbers **I(KT) dont like this sentence - still working on how to word it*.


## Pre-processing and sequencing
DNA extraction was performed as in Stat et al. (2009) with a few modifications. The DNA pellet was washed with 70% ethanol three times, and after the final wash was removed, the remaining ethanol was allowed to evaporate off before the DNA pellet was resuspended in Tris buffer (0.1M pH8).
Cleanup by Amy here
Library Prep
sequencing exact information


ITS2 region - it's annoying, but it's the best we've got right now
PCR and Cleanup - Amy's method of PCR and cleanup
Library Prep - Amy's method of library prep, include Illumina Sequencing information (barcodes, etc)

## Bioinformatics
Sequence QC - boku, then merge with illumina utils, max mismatch=3
Sequence clustering - denovo clustering using UCLUST in QIIME, then compare to reference database to assign taxonomy


## Statistical Analysis 
Alpha diversity of sequence reads, co-occurence?, beta diversity?
Code will be avaible on git hub

# References
<!-- Insert references  -->
```{r, warning=FALSE, message=FALSE, echo=FALSE} 
  write.bibtex(entry=NULL, file="knitcitations.bib", append=FALSE)
```
