---
title: "KI Platygyra Manuscript"
author: "Danielle C Claar, Kristina L Tietjen, Ruth D Gates, Julia K Baum"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
  html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    theme: cerulean
csl: nature.csl
bibliography: references.bib
---

<!-- ## Install and source necessary packages -->
```{r, echo=FALSE}

  # install.packages('knitcitations')
  # install.packages('bibtex')
  # install.packages('qdap')

  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries=FALSE)
  library(bibtex)
  
  # source('http://bioconductor.org/biocLite.R')
  # biocLite('phyloseq')
  # packageVersion('phyloseq')
  library(phyloseq)
  library("qdap") # to be able to use word count
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, incl=FALSE}
  wc("") # to count words
```
  


# Summary
Ocean ecosystems worldwide are threatened by climate change-induced increases in seawater temperatures. Pulse warming events such as El Nino amplify these threats, causing massive losses of coral cover (e.g. 17% of coral reefs during the 1997/98 El Nino). The 2015/16 El Nino is currently the worst on record in terms of severity and longevity, yet despite massive coral mortality, some corals show resilience to this extreme event. Coral resilience is related to many factors, including the structure and flexibility of their internal symbiotic communities. Here we show that the ability of a coral to house a diverse suite of symbionts (*Symbiodinium*) is driven by the identity of the dominant *Symbiodinium* type. Additionally we show that, contrary to current opinion, symbionts present in miniscule abundances (<2%) are indeed important for coral recovery, and furthermore, that corals have the ability to regain symbionts during an extended stress event, providing hope for the future of coral resilience.

# Background *Need to rename using header requirements: The text may contain subheadings (less than six in total) of less than 40 characters (including spaces) each.*
```{r, echo=FALSE, include=FALSE}
citep("Article")
```
Intro goes here. Kiritimati, El Nino, coral bleaching, Platygyra, recovery, resilience
*Articles are typically 3,000 words of text (not including Methods, summary or other sections), beginning with up to 500 words of referenced text expanding on the background to the work, before proceeding to a concise, focused account of the findings, ending with one or two short paragraphs of discussion.*
(corals and climate change) (what about the future of general resilience to climate change) (El Ninos are terrible, this one was the worst so far, and they are getting worse)
(How might corals be resilient - Symbiodinium) (coral bleaching debate) (corals switching symbionts is good or bad?) (corals can regain symbionts ONLY AFTER stress subsides which limits coral recovery when stress increases)
Here, we tagged and sampled the same corals before, during, and immediately after the el nino event, on Kiritimati (something about Kiritimati). We used Illumina sequencing to evaluate changes in symbiodinium community structure coincident with the 2015-2016 major el nino event. The goal was to understand...why the hell these corals survived 10 months of extreme heat stress, and actually got better in the middle of it.
Coral reefs provide vital services, including food security and coastline protection, to coastal communities worldwide. These diverse, vibrant ecosystems are increasingly threatened by human impacts – both at a local scale from water pollution and overfishing, and at a global scale from ocean acidification and climate change. 

Previously, corals have been shown to recover from bleaching only after the external stress (e.g. warming) has subsided.
creating a particularly acute threat to coral reefs, which already live on the edge of their thermal tolerance.

# Findings
*a concise, focused account of the findings, probably <2,000 words*
Results Paragraph 1 (different dominant symbionts drive symbiont diversity capacity, also they didn't switch at commencement of bleaching event)
*Figure about diversity*
Results Paragraph 2 (<2% symbionts are indeed important) - can I calculate some sort of change score for the different types?
*Figure of example sequence abundances, superimposed on coral 99 images*
Results Paragraph 3 (Corals can regain symbionts during an extended stress event)

# Discussion
*one or two short paragraphs of discussion. Probably around 500 wds*
Discussion Paragraph 1 (Methods like NGS are really important for understanding these changes in symbiont diversity, as well as for seeing those low aboundance symbionts)
Discussion Paragraph 2 (What does their recovery tell us about the future of coral reefs?)... Why are Platys so excellent and what does that tell us about when coral resilience is threatened by extreme climatic events?

# Methods
*The Methods section should be written as concisely as possible but should contain all elements necessary to allow interpretation and replication of the results. As a guideline, Methods sections typically do not exceed 3,000 words.*

``````{r, echo=FALSE, warning=FALSE, message=FALSE}
path <- "/Volumes/Seagate_Blue/JuliaBaum_DanielleClaar-32138116_all/"
biom_dir <- "otu_tables/"
biom_file <- paste(path, biom_dir, "KI14FSYM004_otu_table.biom", sep = "")
map_dir <- 
map_file <- paste(path, map_dir, "", sep = "")
ref_dir <- 
ref_file <- 

biom_otu_tax <- import_biom(biom_file, "greengenes")


# Add sample data to the dataset using merge
bmsd <- import_qiime_sample_data(map_file)
class(bmsd)
Bushman <- merge_phyloseq(biom_otu_tax, bmsd)

import_biom(biom, treefilename, refseqfilename, parseFunction=parse_taxonomy_greengenes)
import_biom(rich_dense_biom, treefilename, refseqfilename, parseFunction=parse_taxonomy_greengenes)

```


## Field Information
Kiritimati basics - located in the Central Equatorial Pacific, smack dab in the middle of the Nino 3.4 region (used to quantify el nino presence and strength), human disturbance gradient, bleaching event there (cite bleaching paper here)
Tagging corals and collecting samples - transects, tagging corals, photoing corals, sampling corals, processing samples, storage in Guanidinium
Taxa sampled - platy, favites, favia, etc.
## Pre-processing and sequencing
DNA Extraction - extraction protocol
ITS2 region - it's annoying, but it's the best we've got right now
PCR and Cleanup - Amy's method of PCR and cleanup
Library Prep - Amy's method of library prep, include Illumina Sequencing information (barcodes, etc)
## Post Processing
Sequence QC - boku, then merge with illumina utils, max mismatch=3
Sequence clustering - denovo clustering using UCLUST in QIIME, then compare to reference database to assign taxonomy
Statistical Analysis - alpha diversity of sequence reads, co-occurence?, beta diversity?

# References
<!-- Insert references  -->
```{r, warning=FALSE, message=FALSE, echo=FALSE} 
  write.bibtex(entry=NULL, file="knitcitations.bib", append=FALSE)
```
