---
title: "KI Platygyra Manuscript"
author: "Danielle Claar, Kristina Tietjen, Julia Baum"
date: "19 July 2016"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
  html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    theme: cerulean
csl: mee.csl
bibliography: references.bib
---

```{r, echo=FALSE}
  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries=FALSE)
  library(bibtex)
```

# Abstract

*Abstract goes here. Why are the Platys not bleached anymore*



# Introduction

Intro goes here. Kiritimati, El Nino, coral bleaching, Platygyra, recovery, resilience


# Methods

Kiritimati basics
Tagging corals and collecting samples
DNA Extraction
PCR and Cleanup
Library Prep
Illumina Sequencing (barcodes, region, etc)
Sequence QC
Sequence clustering
Statistical Analysis

## Tables

```{r kable, echo=FALSE}
library(knitr)

x <- rnorm(100)
y <- 2*x + rnorm(100)
out <- lm(y ~ x)
kable(summary(out)$coef, digits=2, caption="This is a GLM summary table.")
```

```{r xtable, results="asis", echo=FALSE, eval=FALSE, message=FALSE}
library(xtable)
tab <- xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
              caption="This is a GLM summary table.")
print(tab, type="latex")
```

## Plots

```{r carDataPlot, echo=FALSE, fig.cap="Relationship between x and y. The solid line is least-squares linear regression.", fig.width=4, fig.height=4}
  plot(x, y)
  abline(out)
```



## Citations

The relationship was first described by
`r citet("10.1111/j.1461-0248.2005.00827.x")`. However, there are also opinions
that the relationship is spurious `r citep("10.1111/j.2041-210x.2012.00264.x")`.
We used R for our calculations `r citep(citation())`, and we used package `knitcitations` `r citep(citation("knitcitations"))` to make the bibliography.


# Results and discussion

Why are Platys so excellent and what does that tell us about when coral resilience is threatened by extreme climatic events?

# References

```{r, warning=FALSE, message=FALSE, echo=FALSE} 
  write.bibtex(file="references.bib")
```




