---
title: "KI Platygyra Manuscript"
author: "Danielle C Claar, Kristina L Tietjen, Ruth D Gates, Julia K Baum"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
  html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    theme: cerulean
csl: nature.csl
bibliography: references.bib
---

<!-- ## Install and source necessary packages -->
```{r, echo=FALSE}

  # install.packages('knitcitations')
  # install.packages('bibtex')
  # install.packages('qdap')

  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries=FALSE)
  library(bibtex)
  
  # source('http://bioconductor.org/biocLite.R')
  # biocLite('phyloseq')
  # packageVersion('phyloseq')
  library(phyloseq)
  library("qdap") # to be able to use word count
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, incl=FALSE}
  wc("") # to count words
```
  
# Summary
Ocean ecosystems worldwide are threatened by climate change-induced increases in seawater temperatures. Pulse warming events such as El Nino amplify these threats, causing massive losses of coral cover (e.g. 17% of coral reefs during the 1997/98 El Nino). The 2015/16 El Nino is currently the worst on record in terms of severity and longevity, yet despite massive coral mortality, some corals show resilience to this extreme event. Coral resilience is related to many factors, including the structure and flexibility of their internal symbiotic communities. Here we show that the ability of a coral to house a diverse suite of symbionts (*Symbiodinium*) is driven by the identity of the dominant *Symbiodinium* type. Additionally we show that, contrary to current opinion, symbionts present in miniscule abundances (<2%) are indeed important for coral recovery, and furthermore, that corals have the ability to regain symbionts during an extended stress event, providing hope for the future of coral resilience.

# Background *rename using header reqs: The text may contain subheadings (<6 in total) of <40 characters (incl spaces) each.*
```{r, echo=FALSE, include=FALSE}
citep("Article")
```

*El Nino is terrible, creating a particularly acute threat to coral reefs, which already live on the edge of their thermal tolerance.* El Niño is the positive phase of the ENSO (El Niño Southern Oscillation). El Niño is a natural climatic event that occurs when surface waters heat up in the equatorial Pacific, causing catastrophic effects on reef ecosystems by disrupting coral symbioses (i.e. coral bleaching). The 2015-2016 El Niño is the first major global event since 1997-1998, and has been declared the third global coral bleaching event by NOAA (cite Reef article). El Ninos are getting worse, threatening coral reefs and endangering the persistence of vital ecosystem services, threatening the food security and coastline protection of coastal communities worldwide. *This el nino is the worst yet* Kiritimati is at the epicenter of the El Niño event, and this region experienced severe thermal anomalies, *what were the anomalies - BAD, cite bleaching paper here*. *Kiritimati, with reefs ranging from shit to pristine, is at the epicenter of the El Nino impact area* *as stressors increase worldwide, all reefs (from shit to pristine) will be increasingly threatened as resilience and recovery are limited by overwhelming environmental stress*

El nino warming threatens coral reefs by disrupting the dynamic symbiosis between coral and their internal symbiotic algae (Symbiodinium). This symbiosis is the foundation of reef ecosystems, and a critical element of reef resilience (van oppen and gates 2006). Corals host a diverse community of Symbiodinium, ranging along a continuum from 'selfish opportunistic symbionts' (e.g. some clade D Symbiodinium) which are better suited to sustained environmental stress than others, to 'intimately evolved symbionts' which provide exceptional amounts of nutrition to their coral host. Thus, although these relationships have developed over evolutionary time, the resilience of the coral symbiome is constantly shaped by dynamic coral-symbiont interactions (stat 2006). *We used to think that bleaching might be good - ABH* *And we do know that some symbio are "better" than others* *And then we said that bleaching is definitely bad* *But at least we do know that it allows changes to occur in the Symbiodinium community structure*

*Regardless it is likely that symbiodinium community composition is important for resilience (Hollie's paper)* *Taxa-specific bleaching is a thing* *When we've seen resilience/recovery before, corals have only been demonstrated to recover if the stress goes away first* Previously, corals have been shown to recover from bleaching only after the external stress (e.g. warming) has subsided. *implying that longer and more frequent stressors spell disaster for reefs worldwide*

Here, we tagged and sampled the same corals before, during, and immediately after the el nino event, on Kiritimati (something about Kiritimati). We used Illumina sequencing to evaluate changes in symbiodinium community structure coincident with the 2015-2016 major el nino event. The goal was to understand...why the hell these corals survived 10 months of extreme heat stress, and actually got better in the middle of it.


# Findings
*a concise, focused account of the findings, probably <2,000 words*
Results Paragraph 1 (different dominant symbionts drive symbiont diversity capacity, also they didn't switch at commencement of bleaching event)
*Figure about diversity*
Results Paragraph 2 (<2% symbionts are indeed important) - can I calculate some sort of change score for the different types?
*Figure of example sequence abundances, superimposed on coral 99 images*
Results Paragraph 3 (Corals can regain symbionts during an extended stress event)

# Discussion
*one or two short paragraphs of discussion. Probably around 500 wds*
Discussion Paragraph 1 (Methods like NGS are really important for understanding these changes in symbiont diversity, as well as for seeing those low aboundance symbionts)
Discussion Paragraph 2 (What does their recovery tell us about the future of coral reefs?)... Why are Platys so excellent and what does that tell us about when coral resilience is threatened by extreme climatic events?

Elucidating the mechanisms underlying changes in coral-symbiont interactions is essential to understanding the ability of the coral symbiome to adapt to the multiple stressors they now face. 

# Methods
*The Methods section should be written as concisely as possible but should contain all elements necessary to allow interpretation and replication of the results. As a guideline, Methods sections typically do not exceed 3,000 words.*

``````{r, echo=FALSE, warning=FALSE, message=FALSE}
path <- "/Volumes/Seagate_Blue/JuliaBaum_DanielleClaar-32138116_all/"
biom_dir <- "otu_tables/"
biom_file <- paste(path, biom_dir, "KI14FSYM004_otu_table.biom", sep = "")
map_dir <- 
map_file <- paste(path, map_dir, "", sep = "")
ref_dir <- 
ref_file <- 

biom_otu_tax <- import_biom(biom_file, "greengenes")


# Add sample data to the dataset using merge
bmsd <- import_qiime_sample_data(map_file)
class(bmsd)
Bushman <- merge_phyloseq(biom_otu_tax, bmsd)

import_biom(biom, treefilename, refseqfilename, parseFunction=parse_taxonomy_greengenes)
import_biom(rich_dense_biom, treefilename, refseqfilename, parseFunction=parse_taxonomy_greengenes)

```


## Field Information
Kiritimati basics - located in the Central Equatorial Pacific, smack dab in the middle of the Nino 3.4 region (used to quantify el nino presence and strength), human disturbance gradient, bleaching event there (cite bleaching paper here)
Tagging corals and collecting samples - transects, tagging corals, photoing corals, sampling corals, processing samples, storage in Guanidinium
Taxa sampled - platy, favites, favia, etc.
## Pre-processing and sequencing
DNA Extraction - extraction protocol
ITS2 region - it's annoying, but it's the best we've got right now
PCR and Cleanup - Amy's method of PCR and cleanup
Library Prep - Amy's method of library prep, include Illumina Sequencing information (barcodes, etc)
## Post Processing
Sequence QC - boku, then merge with illumina utils, max mismatch=3
Sequence clustering - denovo clustering using UCLUST in QIIME, then compare to reference database to assign taxonomy
Statistical Analysis - alpha diversity of sequence reads, co-occurence?, beta diversity?

# References
<!-- Insert references  -->
```{r, warning=FALSE, message=FALSE, echo=FALSE} 
  write.bibtex(entry=NULL, file="knitcitations.bib", append=FALSE)
```
